Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-23 paper.qmd
Updated script to get the projections for UMAP
Updated Figure 11

## First projection
projection <- cbind(
    c(0.18827,-0.28588,-0.20848,-0.78132),
    c(0.71366,0.00788,-0.42630,0.28283))

@@

five_gau_proj_umap_model1 <- projected_df |>
  filter(type == "data") |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        alpha = 0.3) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  xlim(c(-0.18, 0.1)) +
  ylim(c(-0.12, 0.12)) +
  interior_annotation("b", 
                      position = c(0.08, 0.9),
                      cex = 3)

## Second projection
projection <- cbind(
  c(0.58453,-0.38194,0.05395,0.52974),
  c(0.29917,0.29629,-0.76964,-0.03810))

five_gau_proj_umap_model2 <- projected_df |>
  filter(type == "data") |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        alpha = 0.3) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  xlim(c(-0.16, 0.1)) +
  ylim(c(-0.1, 0.1)) +
  interior_annotation("b", 
                      position = c(0.08, 0.9),
                      cex = 3)

@@

```{r}
#| label: fig-five-gau-projs
#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"

# fig-height: 6 
# fig-width: 18

>>>>

# Apply the scaling
scaled_gau_data <- scale_data_manual(training_data_gau |> select(-ID)) |>
  as_tibble()

df_b <- df_b |>
  select(-type)

# Apply the scaling
scaled_gau_data_model <- scale_data_manual(df_b) |>
  as_tibble()

## First projection
projection <- cbind(
    c(0.36638,0.31467,0.63030,0.24465),
    c(0.14672,-0.73585,0.33246,-0.12977))

projected <- as.matrix(scaled_gau_data) %*% projection

projected_df <- projected |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  #dplyr::mutate(type = df_exe$type) |>
  dplyr::mutate(ID = dplyr::row_number())

projected_model <- as.matrix(scaled_gau_data_model) %*% projection

projected_model_df <- projected_model |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  dplyr::mutate(ID = dplyr::row_number())

model_df <- dplyr::left_join(
  distance_df_small_edges |> select(-distance), 
  projected_model_df, 
  by = c("from" = "ID"))

names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        alpha = 0.3) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  xlim(c(-0.45, -0.3)) +
  ylim(c(-0.05, 0.2)) +
  interior_annotation("c", 
                      position = c(0.08, 0.9),
                      cex = 3)

## Second projection
projection <- cbind(
    c(-0.70767,-0.09092,0.15109,-0.39811),
    c(-0.01312,-0.61805,0.44418,0.33305))

@@

five_gau_proj_umap_model2 <- projected_df |>
  filter(type == "data") |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        alpha = 0.3) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  xlim(c(-0.13, 0.13)) +
  ylim(c(-0.13, 0.13)) +
  interior_annotation("d", 
                      position = c(0.08, 0.9),
                      cex = 3)

@@

```{r}
#| label: fig-five-gau-projs
#| fig-height: 12
#| fig-width: 8
#| fig-pos: H
#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
